---
title: "Code for The Social Web"
author: "M.M"
date: "13 March 2020"
output: pdf_document
highlight: tango
header-includes:
   - \usepackage[table,xcdraw]{xcolor}
---

## Data preprocessing
First I extract into 11 files named "output/ratings_sortedXXX-YYY.csv" ratings newer than 01.01.2015 00:00 (in jupyter notebook, python). Then I load them into the dataset in R: 

``` {r}
rating_scale = c(0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0);
sum_ratings <- function(data) {
  result_sum_ratings = c();
  for (i in rating_scale) {
    result_sum_ratings[i * 2] = length(data[data == i]);
  }
  result_sum_ratings
}


preprocess_movie_data_for_sample <- function(data) {
  c(sum_ratings(data[1:(length(data) / 2)]), 
    sum_ratings(data[(length(data) / 2 + 1): length(data)]))
}

preprocess_movie_data <- function(data) {
  random_indices = sample(1:nrow(data), nrow(data) / 2);
  preprocess_sample_1 <- c(data[random_indices,2], data[-random_indices,2]);
  random_indices = sample(1:nrow(data), nrow(data) / 2);
  preprocess_sample_2 <- c(data[random_indices,2], data[-random_indices,2]);
  random_indices = sample(1:nrow(data), nrow(data) / 2);
  preprocess_sample_3 <- c(data[random_indices,2], data[-random_indices,2]);
  c(preprocess_movie_data_for_sample(data[,2]), 
    preprocess_movie_data_for_sample(preprocess_sample_1), 
    preprocess_movie_data_for_sample(preprocess_sample_2), 
    preprocess_movie_data_for_sample(preprocess_sample_3))
}

preprocess_file_data <- function(fileName) {
  data = read.table(fileName, header=TRUE, sep=",");
  movieIds = unique(data[,1]);
  res = matrix(-1,length(movieIds),81);
  counter = 1;
  for (i in movieIds) {
    temp = subset(data,data['movieId'] == i)
    if (nrow(temp) < 50) {
      next
    }
    res[counter,] = c(i, preprocess_movie_data(temp));
    counter = counter + 1;
  }
  res = subset(res, res[,1] != -1);
  res
}
```

__sum_ratings__ function for given ratings produces a vector of 10 values. i-th value represents number of (i/2) ratings in the given data.

__preprocess_movie_data_for_sample__ function for given sample of ratings divides it into 2 halves. As an output vector of 2 x 10 values is produced. First 10 values describes the distribution of ratings in the first half. Next 10 values describes distribution of ratings in the second half.

__preprocess_movie_data__ function for data consisting ratings ordered in time creates a vector of 4 x 20 values. First 20 values describe distribution of ratings in the first half and the second half. Next 3 x 20 values describes distribution of values in the first and second halves for 3 random permutation. These 3 samples are used as a control group. I used 3 samples instead of 1 to avoid unlucky pick. I expect that the median of test results for 3 samples to be >0.05 almost everytime.

__preprocess_file_data__ function for a given fileName reads the data, extracts unique movieIds and then preprocess data from the file. It removes any movie that has less than $50$ ratings from further investigation. As an output it produces matrix of size numberOfMovies x 81. Each row represents a movie that has at least >50 ratings. First column is movieId, then 4 x 20 preprocessed values as described above (4 samples).

## Data testing
For each row we compute 4 times (one time for each sample) chisq_test. This test is most reliable when it comes to measure difference between 2 discrete distributions:

[linked phrase]https://stats.stackexchange.com/questions/298467/are-two-grade-distributions-significantly-different-in-r

[linked phrase]https://rcompanion.org/handbook/H_09.html?fbclid=IwAR0YsHHx7ORNq2qlIrefuW-zRsQgHvjpnXKb6u4qTLLRFeg3ab3RTufWNSQ

``` {r}
library(survival)
library(coin)
chiTest <- function(data, offset) {
  x <- as.table(matrix(data[(offset+1):(offset+20)], nrow=2, byrow=TRUE));
  pvalue(chisq_test(x, scores=list("Var2" = c(1,2,3,4,5,6,7,8,9,10))))
}

test_data <- function(data) {
  movieIds = unique(data[,1]);
  test_results = matrix(-1, length(movieIds), 5);
  for (i in (1:nrow(data))) {
    test_results[i,] = c(data[i,1], 
                         chiTest(data[i,], 1), 
                         chiTest(data[i,], 21), 
                         chiTest(data[i,], 41), 
                         chiTest(data[i,], 61));
  }
  test_results = subset(test_results, test_results[,1] != -1)
}
```

## Data computation and save to file

``` {r}
run_tests <- function() {
  for (i in (2:2)) {
    namePrefix = "output/ratings_sorted";
    inputFilename = paste(namePrefix, 
                          format(((i-1) * 20000), scientific=FALSE), 
                          "-", 
                          format(i*20000, scientific=FALSE), 
                          "__.csv", 
                          sep="");
    outputFilename = paste(namePrefix, 
                           format(((i-1) * 20000), scientific=FALSE), 
                           "-", 
                           format((i*20000), scientific=FALSE), 
                           "__output.csv", 
                           sep="");
    preprocessed_data <- preprocess_file_data(inputFilename);
    tested_data <- test_data(preprocessed_data);
    write.csv(tested_data, outputFilename)
  }
}
```

## Additional Info

To investigate potential points and find any collinearity in our explanatory variables we take a look at plots between all the pairs of explanatory variables.
``` {r, fig.width = 7, fig.height = 6, collapse = TRUE}
library(knitr)
genreMeansData = read.table("output/genres_mean_ratings.csv", header=TRUE, sep=",")
plot(genreMeansData)
kable(genreMeansData)
```

## Marcin's helping functions (just some working space)

``` {r}
gather_results <- function() {
  res = matrix(-1, 0, 5)
  for (i in (1:11)) {
    namePrefix = "output/ratings_sorted";
    inputFilename = paste(namePrefix, 
                          format(((i-1) * 20000), scientific=FALSE), 
                          "-", 
                          format(i*20000, scientific=FALSE), 
                          "__output.csv", 
                          sep="");
    small_res = read.table(inputFilename, header=TRUE, sep=",");
    res = rbind(res, small_res[,2:6]);
  }
  names(res) <- c('movieId', 
                  'pvalue (ordered sample)', 
                  'pvalue (random sample1)', 
                  'pvalue (random sample2)', 
                  'pvalue (random sample3)')
  write.csv(res, "output/ChiTestMovieLens.csv", row.names=FALSE)
  res
}

rating_scale = c(0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0);
sum_ratings <- function(data) {
  result_sum_ratings = c();
  for (i in rating_scale) {
    result_sum_ratings[i * 2] = length(data[data == i]);
  }
  result_sum_ratings
}


preprocess_movie_data_for_sample <- function(data1, data2) {
  c(sum_ratings(data1), sum_ratings(data2))
}

preprocess_movie_data <- function(data, firstYearSize) {
  random_indices = sample(1:nrow(data), firstYearSize);
  preprocess_sample_1 <- c(data[random_indices,2], data[-random_indices,2]);
  random_indices = sample(1:nrow(data), firstYearSize);
  preprocess_sample_2 <- c(data[random_indices,2], data[-random_indices,2]);
  random_indices = sample(1:nrow(data), firstYearSize);
  preprocess_sample_3 <- c(data[random_indices,2], data[-random_indices,2]);
  c(preprocess_movie_data_for_sample(data[1:firstYearSize,2], data[(firstYearSize + 1):nrow(data),2]), 
    preprocess_movie_data_for_sample(preprocess_sample_1[1:firstYearSize], 
                                     preprocess_sample_1[(firstYearSize + 1):length(preprocess_sample_1)]), 
    preprocess_movie_data_for_sample(preprocess_sample_2[1:firstYearSize], 
                                     preprocess_sample_2[(firstYearSize + 1):length(preprocess_sample_2)]), 
    preprocess_movie_data_for_sample(preprocess_sample_3[1:firstYearSize], 
                                     preprocess_sample_3[(firstYearSize + 1):length(preprocess_sample_3)]))
}

preprocess_file_data <- function(fileName) {
  data = read.table(fileName, header=TRUE, sep=",");
  movieIds = unique(data[,1]);
  res = matrix(-1,length(movieIds),81);
  counter = 1;
  oneYearTime = 31622400;
  for (i in movieIds) {
    temp = subset(data,data['movieId'] == i);
    minTime = min(temp['timestamp']);
    maxTime = minTime + oneYearTime;
    temp1 = subset(temp, temp['timestamp'] < maxTime);
    firstYearSize = nrow(temp1);
    if (firstYearSize < 20 || nrow(temp) - firstYearSize < 20) {
      next
    }
    res[counter,] = c(i, preprocess_movie_data(temp, firstYearSize));
    counter = counter + 1;
  }
  res = subset(res, res[,1] != -1);
  res
}


library(survival)
library(coin)
chiTest <- function(data, offset) {
  x <- as.table(matrix(data[(offset+1):(offset+20)], nrow=2, byrow=TRUE));
  pvalue(chisq_test(x, scores=list("Var2" = c(1,2,3,4,5,6,7,8,9,10))))
}

test_data <- function(data) {
  if (nrow(data) < 1) {
    test_results = matrix(-1, 0, 5)
  }
  if (nrow(data) > 0) {
    movieIds = unique(data[,1]);
    test_results = matrix(-1, length(movieIds), 5);
    for (i in (1:nrow(data))) {
      test_results[i,] = c(data[i,1], 
                           chiTest(data[i,], 1), 
                           chiTest(data[i,], 21), 
                           chiTest(data[i,], 41), 
                           chiTest(data[i,], 61));
    }
    test_results = subset(test_results, test_results[,1] != -1)
  }
}

run_tests <- function() {
  for (i in (1:11)) {
    namePrefix = "output/ratings_sorted";
    inputFilename = paste(namePrefix, 
                          format(((i-1) * 20000), scientific=FALSE), 
                          "-", 
                          format(i*20000, scientific=FALSE), 
                          "__.csv", 
                          sep="");
    outputFilename = paste(namePrefix, 
                           format(((i-1) * 20000), scientific=FALSE), 
                           "-", 
                           format((i*20000), scientific=FALSE), 
                           "__output_one_year.csv", 
                           sep="");
    preprocessed_data <- preprocess_file_data(inputFilename);
    print(preprocessed_data)
    tested_data <- test_data(preprocessed_data);
    write.csv(tested_data, outputFilename)
  }
}

gather_results <- function() {
  res = matrix(-1, 0, 5)
  for (i in (1:10)) {
    namePrefix = "output/ratings_sorted";
    inputFilename = paste(namePrefix, 
                          format(((i-1) * 20000), scientific=FALSE), 
                          "-", 
                          format(i*20000, scientific=FALSE), 
                          "__output_one_year.csv", 
                          sep="");
    small_res = read.table(inputFilename, header=TRUE, sep=",");
    res = rbind(res, small_res[,2:6]);
  }
  names(res) <- c('movieId', 
                  'pvalue (ordered sample)', 
                  'pvalue (random sample1)', 
                  'pvalue (random sample2)', 
                  'pvalue (random sample3)')
  # write.csv(res, "output/ChiTestYearFrameMovieLens.csv", row.names=FALSE)
  res
}
```